---
title: "Final Project"
subtitle: "Deli Store Sales Forecasting"
author: "✏️ *Team Name* &nbsp;|&nbsp; ✏️ *Student 1* & ✏️ *Student 2*"
date: today
format:
  revealjs:
    embed-resources: true
    theme: serif
    width: 1600
    height: 900
    margin: 0.03
    slide-number: true
    incremental: false
    scrollable: true
    controls: true
    transition: slide
    footer: "Time Series Forecasting — ITESO | Final Project"
    code-overflow: wrap
    html-math-method: katex
    reference-location: document
execute:
  echo: true
  warning: false
  message: false
highlight-style: atom-one
---

# Setup

## Libraries & Data

```{r}
#| label: setup

library(tidyverse)
library(tsibble)
library(feasts)
library(fable)
library(ggtime)
library(fable.prophet)
```

```{r}
#| label: load-data


load("deli_data.RData")
```


:::{.callout-important}
## Team Name — Read Before Continuing

Throughout this project, every time you see `equipo_X`, replace it with your **Canvas team name**
(e.g., `equipo_1`, `equipo_2`, etc.).

**The R object, the model key inside `model()`, and the file name must all match.**

```r
# Wrong ❌
my_forecast <- ventas_semana |> model(forecast = ETS(total)) |> ...

# Correct ✅
equipo_1_global <- ventas_semana |> model(equipo_1 = ETS(total)) |> ...
```
:::

# Part 1 - Global Store Forecast — Weekly Sales

## Global Store Forecast — Weekly Sales

Train: **Jan 2021 – Dec 2025** &nbsp;|&nbsp; Forecast horizon: **19 weeks** (2026 W01 – W19)

## Overview

```{r}
#| label: p1-overview
#| echo: false

ventas_semana |>
  autoplot(total) +
  labs(
    title    = "Weekly Total Sales — Deli Store GDL",
    subtitle = "Training period: January 2021 – December 2025",
    y        = "MXN",
    x        = NULL
  )
```


## Final Forecast — 2026

:::{.callout-important}
## Use the FULL training data here
Refit your chosen model on all of `ventas_semana` (2021–2025). Do **not** use the internal split.
:::

```{r}
#| label: p1-final-fit
#| cache: true

equipo_X_global <- ventas_semana |>
  model(
    equipo_X = SNAIVE(total)   # ✏️ your final model — team name as key
  ) |>
  forecast(h = 19)             # 19 complete weeks: 2026 W01 – W19

equipo_X_global
```

## Export — Part 1

:::{.callout-important}
## Submission checklist

- [ ] Object `equipo_X_global` is named with your Canvas team name
- [ ] The key inside `model()` matches the object name
- [ ] `h = 19` — do **not** change this
- [ ] File name: `[team_name]_global.RData`
:::

```{r}
#| label: p1-export
#| eval: false

save(equipo_X_global, file = "equipo_X_global.RData")   # ✏️ rename both occurrences
```

# Part 2 - Hierarchical Forecasting — Monthly Sales

## Hierarchical Forecasting — Monthly Sales

Train: **Jan 2021 – Apr 2025** &nbsp;|&nbsp; Forecast horizon: **12 months** (May 2025 – Apr 2026)

## 2.1 Hierarchy Overview

```{r}
#| label: p2-overview
ventas_mes
```

:::{.callout-note}
## Three-level hierarchy
```
Total store
  └── nivel_1  (8 product categories)
        └── nivel_2  (~30 subcategories)
```
- `is_aggregated(nivel_1) & is_aggregated(nivel_2)` → Total
- `!is_aggregated(nivel_1) & is_aggregated(nivel_2)` → Category (nivel_1)
- `!is_aggregated(nivel_1) & !is_aggregated(nivel_2)` → Subcategory (nivel_2) — the bottom level
:::

## Total Sales

```{r}
#| label: p2-total
#| echo: false

ventas_mes |>
  filter(is_aggregated(nivel_1), is_aggregated(nivel_2)) |>
  autoplot(total) +
  labs(
    title    = "Total Monthly Sales — Deli Store GDL",
    subtitle = "Training period: January 2021 – April 2025",
    y        = "MXN", x = NULL
  )
```


## Sales by Category

```{r}
#| label: p2-nivel1
#| fig-height: 6
#| echo: false

ventas_mes |>
  filter(!is_aggregated(nivel_1), is_aggregated(nivel_2)) |>
  autoplot(total) +
  facet_wrap(vars(nivel_1), scales = "free_y", ncol = 2) +
  theme(legend.position = "none") +
  labs(
    title = "Monthly Sales by Category (nivel_1)",
    y     = "MXN", x = NULL
  )
```


## Subcategory Drill-Down

```{r}
#| label: p2-nivel2
#| echo: false

ventas_mes |>
  filter(
    !is_aggregated(nivel_2),
    nivel_1 == "Carnes y Embutidos"   # ✏️ explore 2-3 categories of your choice
  ) |>
  autoplot(total) +
  facet_wrap(vars(nivel_2), scales = "free_y") +
  theme(legend.position = "none") +
  labs(
    title = "Subcategories — Carnes y Embutidos",   # ✏️ update title
    y     = "MXN", x = NULL
  )
```


:::{.callout-note}
## Available reconciliation methods

| Method | Description |
|--------|-------------|
| `bottom_up(base)` | Aggregate bottom-level forecasts upward |
| `top_down(base)` | Disaggregate top-level forecasts downward |
| `min_trace(method = "ols")` | Minimise trace of forecast error covariance (OLS weights) |
| `min_trace(method = "wls_struct")` | WLS with structural scaling |
| `min_trace(method = "mint_shrink")` | MinT with shrinkage estimator — recommended when many series |
:::



## Export — Part 2

:::{.callout-important}
## Submission checklist

- [ ] Object `equipo_X_hier` is named with your Canvas team name
- [ ] The reconciliation step key matches the object name (the `.model` column value)
- [ ] `h = 12` — do **not** change this
- [ ] File name: `[team_name]_hier.RData`
:::

```{r}
#| label: p2-export
#| eval: false

save(equipo_X_hier, file = "equipo_X_hier.RData")   # ✏️ rename both occurrences
```


